# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))

```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
daly_burden_ncd <- data_ncd %>% 
  filter(Entity %in% c("Kenya", "South Africa", "Zimbabwe")) %>% 
  rename(daly = 4)

# You should pivot the data to show each country in a separate column.
daly_wide_ncd <- daly_burden_ncd %>% 
  pivot_wider(
    names_from = Entity,
    values_from = daly,
    id_cols = Year)

# Use kable() from the knitr package to render the table.
kable(daly_wide_ncd)
```

## Plot Showing Trends in NCD Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.


daly_trend_plot <- daly_burden_ncd %>% 
  ggplot(mapping = aes(x = Year,
                       y = daly,
                       colour = Entity)) +
           geom_line()+
  labs(title = "Trends in NCD Burden Over Time",
       x = "Year",
       y = "DALY (Disability Adjusted Life Years)") +
    theme_classic() +
   theme(
        plot.title = element_text(color = "black", size = 20, face = "bold"),
        plot.subtitle = element_text(color = "black", size = 10, face = "italic"), 
        axis.title.x = element_text(color = "black", size = 12, face = "bold"),
        axis.title.y = element_text(color = "black", size = 12, face = "bold"),
        axis.text.x = element_text(hjust = 0.5, vjust=0.5),
        axis.text.y = element_text(size = 10))

daly_trend_plot
```

## Summary of NCD Burden Findings

Based on the chart, the NCD burden in South Africa was higher compared to Kenya and Zimbabwe. Kenya had a lower DALY trend reaching a plateau around 2007-2011. For South Africa, there is a notable spike just before the year 2000, before the trend started to decline whereas the trend in Zimbabwe increased exponentially around 1997 to 2010 before starting to decrease.
